---
title: "InfIntE"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What is InfIntE?

InfIntE stands for Inference of Interactions using Explainable machine learning. This package uses abundance data to directly infer ecological interactions using PyGol, an Aductive/Inductive logic program. The interactions inferred are directly classified by its interaction type. 

# Table of contents
1. [Installation](#installation)
2. [Example Data](#test)

## Installation <a name="installation"></a>

InfInt and required packages are installed using devtools

```{r }
library(devtools)
if(!"InfIntE" %in% rownames(installed.packages())){
  install_github("didacb/InfIntE")
}
library(InfIntE)
```

One of the steps of the interaction inference is the abduction. Abduction is performed using PyGol. PyGol is written in c. To compile PyGol and obtain the functions for abduction run: 

```{r}
load_PyGol()
```

## Example Data <a name="test"></a>

```{r}
library(phyloseq)
phylos<- readRDS("~/Downloads/BCM_ITS_wheat_phyloseq_filtered.RDS")
pr<- sample_data(phylos)$Date == "03_18" & sample_data(phylos)$Specie == "wheat" & sample_data(phylos)$Tissue == "G"  
new.phyl<- prune_samples(pr, phylos)
saveRDS(new.phyl, "~/wheat_OTU_data.RDS")
 
a<- infinte(otu_tb = otu_table(new.phyl, taxa_are_rows = T))
```
%%### Sub paragraph <a name="subparagraph1"></a>
%%%This is a sub paragraph, formatted in heading 3 style

