---
title: "InfIntE"
author: "Didac Barroso-Bergada"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What is InfIntE?

InfIntE stands for Inference of Interactions using Explainable machine learning. This package uses abundance data to directly infer ecological interactions using PyGol, an Aductive/Inductive logic program. The interactions inferred are directly classified by its interaction type. 

# Table of contents
1. [Installation](#installation)
2. [Some paragraph](#paragraph1)
    1. [Sub paragraph](#subparagraph1)
3. [Another paragraph](#paragraph2)

## Installation <a name="installation"></a>

InfInt and its dependences are installed using devtools

```{r}
library(devtools)
install_github("didacb/InfIntE", auth_token = "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")
library(InfIntE)
```

```{r}
load_PyGol()
```

## Some paragraph <a name="paragraph1"></a>
The first paragraph text

### Sub paragraph <a name="subparagraph1"></a>
This is a sub paragraph, formatted in heading 3 style

## Another paragraph <a name="paragra


```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
